---
title: "Palmer Penguins: Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---


```{r,include=FALSE}
# LOAD PACKAGES

rm(list = ls())

library(flexdashboard)
library(shinyWidgets)
library(shiny)
library(tidyverse)
library(plotly)
library(DT)
library(palmerpenguins)
library(htmltools)




# READ IN DATA



dat <- palmerpenguins::penguins





# DATA PROCESSING I: WIDGETS



## WIDGETS FOR CATEGORIES: ISLAND AND SPECIES



species <-unique(dat$species)










```





<b>Penguin Species</b> {data-orientation=rows}
=====================================  


Sidebar {.sidebar}
-------------------------------------


```{r}

selectInput("species", label = HTML('<b style="color:#fffff;
               font-size: 20px;
               ">Select Penguin Species:</b>'), 
    choices = species, 
    selected = "Chinstrap")


radioButtons("plotType", HTML('<b style="color:#fffff;
               font-size: 20px;
               ">Select Plot or Table:</b>'),
             choices = c("Histogram Chart","Data Table"),
             selected = "Histogram Chart")


```






Column1 
-------------------------------------


### <b>Total Count</b>

```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(species %in% input$species)
  
  penguins <- count(d2)
  
  valueBox(penguins, 
           icon = "fa fa-user-circle")
  
})



```





### <b>Penguin Species</b>

```{r}


renderValueBox({
  
  d2 <- dat %>%
    filter(species %in% input$species) 
  
  penguins <- unique(d2$species)
  
  valueBox(penguins, 
           icon = "fa fa-users")
  
})





```






### <b>Avg. Body Mass<br>(Grams)</b>


```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(species %in% input$species)  
  
  penguins <- ceiling(mean(d2$body_mass_g,na.rm=T))
  
  valueBox(penguins, 
           icon = "fa fa-moon-o")
  
})










```







Row2
-------------------------------------

### <b>Palmer Penguins Table</b>

```{r}

uiOutput("dynamic")

output$dynamic <- renderUI({ 
  if (input$plotType == "Histogram Chart") plotOutput("plot")
  else if (input$plotType == "Data Table") dataTableOutput("table")
})

output$plot <- renderPlot({ 
  
  
  d2 <- dat %>%
    filter(species %in% input$species)
  
  ggplot(d2) +
    aes(x = body_mass_g) +
    geom_histogram(bins = 30L, fill = "#112446") +
    theme_minimal()
  
  
  })

output$table <- renderDT({ 
  
  dat %>%
    filter(species %in% input$species)
  
  
  
  
  })



  
  
 

  
```




<b>Statistics</b> {data-orientation=rows}
=========================================


Column 1
-----------------------------------------

### <b>Bill length vs. depth</b>

```{r}


renderPlotly({
  
  
  bill_len_dep <- ggplot(data = penguins,
                         aes(x = bill_length_mm,
                             y = bill_depth_mm,
                             group = species)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, aes(color = species)) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Penguin bill dimensions",
       subtitle = "Bill length and depth for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Bill length (mm)",
       y = "Bill depth (mm)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.85, 0.15),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

ggplotly(bill_len_dep)
  
  
})

```


### <b>Flipper Histogram</b>


```{r}

renderPlotly({
  
  
  
flipper_hist <- ggplot(data = penguins, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species), 
                 alpha = 0.5, 
                 position = "identity") +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) +
  labs(x = "Flipper length (mm)",
       y = "Frequency",
       title = "Penguin flipper lengths")

ggplotly(flipper_hist)
  
  
  
})


```



<b>About</b>
=====================


Row 
-------------------------------------

### <b>About this Dashboard</b>





These data were collected from 2007 - 2009 by Dr. Kristen Gorman with the Palmer Station Long Term Ecological Research Program, part of the US Long Term Ecological Research Network. The data were imported directly from the Environmental Data Initiative (EDI) Data Portal, and are available for use by CC0 license (“No Rights Reserved”) in accordance with the Palmer Station Data Policy.


**Data Dictionary:**

 species: a factor denoting penguin species (Adélie, Chinstrap and Gentoo)
 
 island: a factor denoting island in Palmer Archipelago, Antarctica (Biscoe, Dream or Torgersen)
 
 bill_length_mm: a number denoting bill length (millimeters)
 
 bill_depth_mm: a number denoting bill depth (millimeters)
 
 flipper_length_mm: an integer denoting flipper length (millimeters)
 
 body_mass_g: an integer denoting body mass (grams)
 
 sex: a factor denoting penguin sex (female, male)
 
 year: an integer denoting the study year (2007, 2008, or 2009)






### <b>palmerpenguins package</b>

Allison Horst. Author, Maintainer. {https://www.allisonhorst.com/}

Alison Hill. Author. {https://www.apreshill.com/}

Kristen Gormen. Author. {https://www.uaf.edu/cfos/people/faculty/detail/kristen-gorman.php}




Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0. https://allisonhorst.github.io/palmerpenguins/. doi: 10.5281/zenodo.3960218.


```{r}
citation("palmerpenguins")
```




```{css,echo = FALSE}


/* Palmer Penguins Shiny Dashboard */

/* Format Chart Headers */

.chart-title {
    border-bottom: 1px solid #dfdfdf;
    font-size: 14px;
    font-weight: 300;
    padding: 7px 10px 4px;
    color: white;
    font-family: inherit;
    background-color: purple;
}


/* Format Navigation Bar for Dashboard */
.navbar-inverse {
    background-color: purple;
    border-color: purple;
}


.navbar, [data-toggle=tab], .navbar-brand  {
background-color:purple!important;
boarder-color: #ffffff!important;}




/* Sidebar Format */
.section.sidebar {
    top: 51px;
    background-color: darkorange;
    color: #ffffff;
}


/* Format the valueboxes */


/* This will make the the text color white for the valueboxes */

.value-box .value {
    background-color: darkorange;
    color: #ffffff;
}
.value-box .caption {
    background-color: darkorange;
    color: #ffffff;
}

.value-box-primary {
    background-color: darkorange;
    color: #ffffff;
}











```
